An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-07-12 neurips-19/poly-embed.tex
Edits after convo with raf. bo@boulder

  $L(\cdot;p)$ is a convex polyhedral function, so it is the pointwise maximum over finitely many linear functions.
  Construct the convex polyhedral function $\hat{L}(\cdot;p)$ by dropping from the max those linear functions that are never equal to $L$ for any $u^* \in U$.
  We have $\hat{L}(u^*;p) = L(u^*;p)$ for all $u^* \in U$ and $\hat{L}(u;p) \leq L(u;p)$ for all $u \not\in U$.
  Now $\hat{L}$ is the max over a finite number of linear functions.
  Each such function $f$ with gradient $\nabla f$ is equal to $\hat{L}$ above a closed, convex cell in the power diagram formed by projecting $\hat{L}(\cdot;p)$.
  If $f$ has nonzero gradient, its cell overlaps with $U$ exactly at some face of $U$.
  We will prove that there exists $c_f > 0$ such that, for all $u$ in the corresponding cell of the power diagram,
    \[ \hat{L}(u;p) \geq L(u^*;p) + c_f \cdot d(U,u) . \]
  We will then repeat this for the finitely many $f$ with nonzero gradient (which covers all points $u \not\in U$) and take the minimum to obtain the desired $c > 0$.
  This will complete the proof.

  Consider the set of unit vectors $\{v \in \reals^d : \|v\|=1\}$ and the set of $u^*$ on the boundary of $U$.
  For all pairs $u^*,v$ such that $v$ exposes $u^*$, let
    \[ G_{u^*,v} = \left\{ u^* + \beta v : \beta \geq 0 \right\} . \]
  Note that $\{ u : u \not\in U\} \subseteq \cup_{u^*,v} G_{u^*,v}$.
  We proceed as follows:
  \begin{enumerate}
    \item Each $G_{u^*,v}$ can be associated with at least one of the finitely many linear functions $f$ as follows: $f(u) = \hat{L}(u;p)$ for all $u \in G_{u^*,v}$.
    \item There exists a constant $c_{u^*,v} > 0$ such that $f(u) = L(u^*;p) + c_{u^*,v} \cdot d(U,u)$ for all $u \in G_{u^*,v}$.
    \item The set of $(u^*,v)$ associated with $f$ is compact (closed and bounded), and the map $(u^*,v) \mapsto c_{u^*,v}$ is continuous.
    \item Hence, there is an infimum $c > 0$ such that $f(u) \geq L(u^*;p) + c \cdot d(U,u)$ for all $u$ in all $G_{u^*,v}$ associated with $f$.
    \item Taking the minimum such $c$ over these finitely many $f$, we obtain $L(u;p) \geq \hat{L}(u;p) \geq inf_{u^*} L(u^*;p) + c \cdot d(U,u)$ for all $u \not\in U$.
  \end{enumerate}

  (1) Follows because $G_{u^*,v}$ is contained in a cell of the power diagram, so it lies below the function where $\hat{L}(\cdot;p) = f(\cdot)$. \bo{Should have more justification!}

  (2) Let $\nabla f$ denote the gradient of the linear function $f$.
  Note that $f(u^*) = \hat{L}(u^*;p) = L(u^*;p)$.
  So we can write
  \begin{align*}
    f(u) &= f(u^*) + (\nabla f)\cdot (d(u^*,u) v)  \\
         &= f(u^*) + c_{u^*,v} \cdot d(U,u)
  \end{align*}
  where $c_{u^*,v} = (\nabla f)\cdot v$ and we used that $d(u^*,u) = d(U,u)$ by Lemma \ref{lemma:exposed-shortest}.
  We must have $c_{u^*,v} > 0$ because the set $U$ minimizes $L(\cdot;p)$, which is upper-bounded by $f$, so $f(u) > f(u^*) = L(u^*;p)$.

  (3) \bo{This one needs more details}
  This follows because the set of $u^*$ such that $f(u^*) = \hat{L}(u^*;p)$ is closed, and for fixed $u^*$, the set of $v$ such that $G_{u^*,v}$ is in the cell of the power diagram is also closed.
  The set is also bounded \bo{is there a corner case to consider here?} and the map $(u^*,v) \mapsto (\nabla f)\cdot v =: c_{u^*,v}$ is continuous for fixed $f$ (it is linear in $v$ only).

>>>>

  $L(\cdot;p)$ is a convex polyhedral function, so it is the pointwise maximum over finitely many linear functions.
  Recall that $\risk{L}(p) = \min_{u} L(u;p)$, the Bayes risk.
  Construct the convex polyhedral function $\hat{L}(\cdot;p)$ by dropping from the max those linear functions that are never equal to $L$ for any $u^* \in U$.
  We have $\hat{L}(u^*;p) = \risk{L}(p)$ for all $u^* \in U$ and $\hat{L}(u;p) \leq L(u;p)$ for all $u \not\in U$.
  Now $\hat{L}$ is the max over a finite number of linear functions.
  Each such function $f$ with gradient $\nabla f$ is equal to $\hat{L}$ above a closed, convex cell in the power diagram formed by projecting $\hat{L}(\cdot;p)$.
  If $f$ has nonzero gradient, its cell overlaps with $U$ exactly at some face of $U$.
  We will prove that there exists $c_f > 0$ such that, for all $u$ in the corresponding cell of the power diagram,
    \[ \hat{L}(u;p) \geq L(u^*;p) + c_f \cdot d(U,u) . \]
  We will then repeat this for the finitely many $f$ with nonzero gradient (which covers all points $u \not\in U$) and take the minimum to obtain the desired $c > 0$.
  This will complete the proof.

  Consider the set of unit vectors $\{v \in \reals^d : \|v\|=1\}$ and the set of $u^*$ on the boundary of $U$.
  For all pairs $u^*,v$ such that $v$ exposes $u^*$, let
    \[ G_{u^*,v} = \left\{ u^* + \beta v : \beta \geq 0 \right\} . \]
  %Note that $\{ u : u \not\in U\} \subseteq \cup_{u^*,v} G_{u^*,v}$.
  We will prove the following steps:
  \begin{enumerate}
    \item Each $G_{u^*,v}$ can be associated with at least one of the finitely many linear functions $f$ as follows: $f(u) = \hat{L}(u;p)$ for all $u \in G_{u^*,v}$.
    \item There exists a constant $c_{v} > 0$ such that $L(u) = \risk{L}(p) + c_{v} \cdot d(U,u)$ for all $u \in G_{u^*,v}$ and all $u^*$ exposed by $v$.
    \item The set of unit vectors $v$ that can be associated with $f$ is compact (closed and bounded), and the map $v \mapsto c_{v}$ is continuous on this set.
    \item Hence, there is an infimum $c > 0$ such that $f(u) \geq \risk{L}(p) + c \cdot d(U,u)$ for all $u$ in all $G_{u^*,v}$ associated with $f$.
    \item Taking the minimum such $c$ over these finitely many $f$, we obtain $L(u;p) \geq \hat{L}(u;p) \geq \risk{L}(p) + c \cdot d(U,u)$ for all $u \not\in U$.
  \end{enumerate}

  (1) Follows because $G_{u^*,v}$ is contained in a cell of the power diagram, which is a set of points where $\hat{L}(\cdot;p) = f(\cdot)$. \bo{Should have more justification.}

  (2) Let $\nabla f$ denote the gradient of the linear function $f$.
  Note that because $u^*$ is on the boundary of $U$, we have $f(u^*) = \hat{L}(u^*;p) = \risk{L}(p)$.
  So we can write, using Lemma \ref{lemma:exposed-shortest},
  \begin{align*}
    f(u) &= f(u^*) + (\nabla f)\cdot (u - u^*)  \\
    f(u) &= f(u^*) + (\nabla f)\cdot (d(u^*,u) v)  \\
         &= \risk{L}(p) + c_{v} \cdot d(U,u)  \\
  \end{align*}
  where $c_{v} = (\nabla f)\cdot v$.
  We must have $c_{v} > 0$ because the set $U$ minimizes $L(\cdot;p)$, so $f(u) > f(u^*) = \risk{L}(p)$.

  (3) The set of $u^*$ where $f(u^*) = \risk{L}(p)$ is the intersection of two cells of the power diagram, i.e. the boundary of $U$ with the cell associated with $f$.
  Hence it is closed and convex.
  The set of unit vectors $v$ that expose a closed, convex set of boundary points is also closed and convex.
  \bo{That should be justified/rechecked.}
  It is also bounded, hence compact.
